Discrep_Patterns

### 2024-11-04 misc/discrep.qmd
discrep update

s2_long %>%
  filter(appliance != "TOTAL", plan == "plan1") %>%
  group_by(appliance, rounded, state) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, family, rounded, state,color_flag) %>%
  summarise(n = n(), .groups = "drop")

>>>>

correct_values_s2 <- tribble(
  ~state, ~appliance, ~family_exact, ~family_rounded, ~state_avg,
  # Texas values (Smith Family)
  "Texas", "Cooling", 6573, 7000, 4249,
  "Texas", "Heating", 6118, 6000, 5099,
  "Texas", "Water Heating", 5257, 5000, 4396,
  "Texas", "Refrigerator", 2639, 3000, 1318,
  "Texas", "Other Appliance", 8459, 8000, 7883,
  "Texas", "Total kWh", 29046, 29000, 22945,

  # California values (Adams Family)
  "California", "Cooling", 2581, 3000, 1289,
  "California", "Heating", 6157, 6000, 5597,
  "California", "Water Heating", 5061, 5000, 4601,
  "California", "Refrigerator", 1266, 1000, 1055,
  "California", "Other Appliance", 7608, 8000, 6916,
  "California", "Total kWh", 22673, 23000, 19458,

  # Colorado values (Wells Family)
  "Colorado", "Cooling", 697, 1000, 498,
  "Colorado", "Heating", 18052, 18000, 16411,
  "Colorado", "Water Heating", 11667, 12000, 5832,
  "Colorado", "Refrigerator", 1370, 1000, 1142,
  "Colorado", "Other Appliance", 7982, 8000, 6652,
  "Colorado", "Total kWh", 39768, 40000, 30535,

  # Massachusetts values (Davis Family)
  "Massachusetts", "Cooling", 419, 0, 322,
  "Massachusetts", "Heating", 26751, 27000, 19108,
  "Massachusetts", "Water Heating", 10543, 11000, 5070,
  "Massachusetts", "Refrigerator", 1230, 1000, 1025,
  "Massachusetts", "Other Appliance", 7350, 7000, 6682,
  "Massachusetts", "Total kWh", 46293, 46000, 32207
) %>%
  rowwise() %>%
  mutate(
    state_abrv = case_when(
      state == "Texas" ~ "TEX",
      state == "California" ~ "CAL",
      state == "Colorado" ~ "COL",
      state == "Massachusetts" ~ "MASS"
    ),
    family_name = case_when(
      state == "Texas" ~ "Smith Family",
      state == "California" ~ "Adams Family",
      state == "Colorado" ~ "Wells Family",
      state == "Massachusetts" ~ "Davis Family"
    ))

s2_long %>%
  filter( plan == "plan1") %>%
  group_by(appliance, rounded, state,pct_goal) %>%
  mutate(
    distinct_families = n_distinct(family),
    distinct_state_avg = n_distinct(state_avg),
  ) %>% 
  filter(distinct_families > 1 | distinct_state_avg > 1)  |> 
   group_by(appliance, rounded, state,pct_goal, family) |>
   mutate(family_count = n()) |>
   group_by(appliance, rounded, state,pct_goal, state_avg) |>
    mutate(state_avg_count = n()) |>
  distinct(rounded,state,refClass,appliance,family,state_avg,state_avg_count,family_count) |> 
     rowwise() |> 
  mutate(
    fam_in = if (rounded == "Rounded") {
      family %in% correct_values_s2$family_rounded
    } else {
      family %in% correct_values_s2$family_exact
    },
    state_in = state_avg %in% correct_values_s2$state_avg
  ) |> arrange(rounded,refClass,state,appliance,family_count) |> 
  relocate(rounded,pct_goal,refClass,state,appliance,family,family_count,state_avg,state_avg_count,fam_in,state_in) |> 
  kable()

  # ungroup() %>%
  # group_by(appliance, family, rounded, state) %>%
  # summarise(n = n(), .groups = "drop")
